---
title: "ISLR2 Exercise 5.9 — Bootstrap on the Boston Housing Data"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    theme: readable
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

# Introduction

This document works through Exercise 5.9 from *An Introduction to Statistical
Learning with Applications in R* (James, Witten, Hastie, Tibshirani). We use the
`Boston` housing data set from the `ISLR2` package and apply the bootstrap to
estimate standard errors for the mean, median, and tenth percentile of `medv`
(median value of owner-occupied homes, in thousands of dollars).

> **Note.** Bootstrap results vary slightly with the random seed and package
> version, so your exact standard errors may differ by a hundredth or two from
> the numbers shown in the comments below.

```{r libraries}
library(ISLR2)
library(boot)

# Reproducibility
set.seed(1)

# Attach for convenience (medv can then be referenced directly)
attach(Boston)
```

---

# (a) Estimate of the population mean

The natural estimate of the population mean is the sample mean.

```{r part-a}
mu.hat <- mean(medv)
mu.hat
```

So $\hat{\mu} \approx 22.53$ (in thousands of dollars).

---

# (b) Standard error of $\hat{\mu}$ (formula)

Using $\text{SE}(\hat{\mu}) = s / \sqrt{n}$:

```{r part-b}
n <- length(medv)
se.mu.hat <- sd(medv) / sqrt(n)
se.mu.hat
```

**Interpretation.** $\text{SE}(\hat{\mu}) \approx 0.409$. If we repeatedly drew
samples of size `r n` from the population and computed the mean each time, those
sample means would vary around the true mean with a standard deviation of about
0.41. It quantifies the uncertainty in $\hat{\mu}$ as an estimate of the true
population mean.

---

# (c) Standard error of $\hat{\mu}$ (bootstrap)

```{r part-c}
boot.fn <- function(data, index) {
  mean(data[index])
}

set.seed(1)
boot.mean <- boot(medv, boot.fn, R = 1000)
boot.mean
```

The bootstrap standard error (≈ 0.41) is essentially identical to the
formula-based estimate from part (b). This is reassuring: the bootstrap recovers
the analytic answer without relying on the formula.

```{r part-c-store}
se.boot.mean <- sd(boot.mean$t)
se.boot.mean
```

---

# (d) 95% confidence interval

Using the approximation $[\hat{\mu} - 2\,\text{SE}, \; \hat{\mu} + 2\,\text{SE}]$
with the bootstrap standard error:

```{r part-d}
ci.boot <- c(mu.hat - 2 * se.boot.mean,
             mu.hat + 2 * se.boot.mean)
ci.boot

# Compare with the t-based interval
t.test(medv)
```

The two intervals are nearly identical (roughly `[21.71, 23.35]` from the
bootstrap vs `[21.73, 23.34]` from `t.test`). The slight difference comes from
using the round factor 2 rather than the exact $t$-quantile (≈ 1.965), plus
bootstrap sampling variation.

---

# (e) Estimate of the population median

```{r part-e}
med.hat <- median(medv)
med.hat
```

So $\hat{\mu}_{\text{med}} = 21.2$.

---

# (f) Standard error of the median (bootstrap)

There is no simple closed-form standard error for the median, so the bootstrap
is the practical tool.

```{r part-f}
boot.fn.med <- function(data, index) {
  median(data[index])
}

set.seed(1)
boot.med <- boot(medv, boot.fn.med, R = 1000)
boot.med
```

**Comment.** $\text{SE}(\hat{\mu}_{\text{med}}) \approx 0.38$. The median (21.2)
has a small standard error relative to its value — smaller even than the SE of
the mean — so we have estimated it fairly precisely. This is valuable precisely
because no easy formula exists.

---

# (g) Tenth percentile

```{r part-g}
mu.hat.0.1 <- quantile(medv, 0.1)
mu.hat.0.1
```

So $\hat{\mu}_{0.1} = 12.75$.

---

# (h) Standard error of the tenth percentile (bootstrap)

```{r part-h}
boot.fn.0.1 <- function(data, index) {
  quantile(data[index], 0.1)
}

set.seed(1)
boot.0.1 <- boot(medv, boot.fn.0.1, R = 1000)
boot.0.1
```

**Comment.** $\text{SE}(\hat{\mu}_{0.1}) \approx 0.48$. This is larger than the
standard error of the mean or median. That is expected: tail quantiles are
estimated from fewer effective observations near that point, so they are
inherently noisier. Still, 0.48 is small relative to the estimate of 12.75, so
the tenth percentile is estimated with reasonable precision.

---

# Summary

```{r summary-table, echo=FALSE}
summary_df <- data.frame(
  Quantity = c("Mean (a)", "Median (e)", "10th percentile (g)"),
  Estimate = c(round(mu.hat, 3),
               round(med.hat, 3),
               round(as.numeric(mu.hat.0.1), 3)),
  `Bootstrap SE` = c(round(se.boot.mean, 3),
                     round(sd(boot.med$t), 3),
                     round(sd(boot.0.1$t), 3)),
  check.names = FALSE
)
knitr::kable(summary_df, caption = "Point estimates and bootstrap standard errors for medv.")
```

```{r cleanup, include=FALSE}
detach(Boston)
```
